Dynamics of Anthropomorphism in Human-Robot Interaction

### 2014-11-07 paper.tex
Large simplification of the paper (now 23 pages)

\usepackage{graphicx}
\graphicspath{{figs/}}
\usepackage{tikz}
\usetikzlibrary{shapes}
\usetikzlibrary{svg.path}
%
\usepackage{amsmath} 
\usepackage{mathptmx}      % use Times fonts if available on your TeX system
%
\usepackage{color, soul}
\usepackage{url}
\usepackage{multirow}
\usepackage{array}
\usepackage{fixltx2e}
\usepackage{textcomp}

@@

%\paragraph{Two hypotheses}
Our research question also translates to two
hypotheses, which structure the reflections of this article:

\begin{enumerate} 
\item ~A person's tendency to anthropomorphize a robot
    is based on several factors, including (but not limited to) (1) the
    design of the robot, (2) the personal characteristics of the human user,
    and (3) the context and situation of use. For instance, a social context
    and purpose of the robot increase anthropomorphism: \textit{anthropomorphism
    as a multi-factor phenomenon}.

\item ~A person's tendency to anthropomorphize a robot
    evolves in a non-linear way over time and with growing experience that the
    person has with the robot. For instance, a more experienced user will
    perceive the robot as more familiar and predictable, and in turn
    anthropomorphizes the robot less: \textit{anthropomorphism as a dynamic
    phenomenon}.

@@

The remainder of this article is organized as follows.  According to the three
perspectives that we take, we first (in Section \ref{sec:anthropomorphism}) aim
to establish a common ground in the understanding of anthropomorphism. We review
literature from several relevant domains and integrate the different
explanations into an understanding of anthropomorphism as a multi-factor
phenomenon.  Based on this, Section~\ref{sec:our-ideas} introduces our initial
model of the \textit{Dynamics of Anthropomorphism}. This model describes the
variances and dynamics of people's tendency to anthropomorphize an artificial
agent, such as a robot.  Then Section \ref{sec:cognition-neuroscience} presents
a cognitive and neuroscience perspective on anthropomorphism to consolidate our
considerations and the proposed model.  This is followed by a discussion about
the limitations of our approach (Section \ref{sec:discussion}) and the existing
and alternative ways to measure anthropomorphism in human-robot interactions.
Finally, Section \ref{sec:conclusion} concludes this article and offers starting
points for further research on anthropomorphism and the effect of the human-like
design approach in robotics.

@@

\section{Towards a unified understanding: What is anthropomorphism?}
\label{sec:definitions}

Literature on anthropomorphism is quite diverse. There exist different and
partly conflicting definitions of the term anthropomorphism, and that in
general, the \textit{terminology} is not clear. Probably as a result of this,
there also exist different understandings of the \textit{phenomenon} that the
term ``anthropomorphism'' describes.  In this section we will first give an
overview on the diversity of terms and then try to establish a common ground in
the understanding of anthropomorphism as a psychological phenomenon.

@@

\begin{figure}
    \centering
    \includegraphics[width=0.75\columnwidth]{anthropo.pdf}
    \caption{Anthropomorphism is a psychological phenomenon which arises in a
        social context, during an interaction between a human agent and a
        non-human agent. One can take two main perspectives when trying to
        explain anthropomorphism: a human-centered and an artifact-centered
        perspective (grey triangles). Several external factors influence the
        anthropomorphism phenomenon. Person- and artifact-related characteristics,
        but also related to the context of use as well as the actual time spent
        interacting with the robot (interaction history). [SEE
        FIG~\ref{fig:anthropofig-alt} FOR AN ALTERNATIVE FIGURE]}

@@

\section{Why do humans anthropomorphize?}
\label{sec:anthropomorphism}

Anthropomorphism is an interesting phenomenon because it tells us something
about the relationship between the human who anthropomorphizes, and the
non-human agent which is anthropomorphized. Though commonly observed, it still
appears controversial, at the same time, that humans tend to \textit{``see human
in the non- human''} \citep{epley_seeing_2007}.

As anthropomorphism arises in an interaction between a human and a non-human
agent \citep{persson_anthropomorphism_2000}, one can also take two main
\textit{perspectives} when trying to explain why humans anthropomorphize
\citep{lee_human_2005}: an artifact-centered perspective and a human-centered
perspective (see grey triangles in Figure \ref{fig:anthropofig}). It has to be
kept in mind that both perspectives are inseparable; they apply at the same
time, and we would like to present them here just as two different viewpoints on
the same phenomenon.

In the following we will only briefly present the artifact-centered perspective
to explain anthropomorphism. A review of anthropomorphic design of robots can
also be found in \cite{fink_anthropomorphism_2012}. Then, we provide a more
detailed review and synthesis of literature that applies a human-centered
perspective to explain anthropomorphism. Concretely, we integrate theories from
developmental psychology, cognitive science / neuroscience, and social
psychology to shed light on the question why humans anthropomorphize.

@@

\item ~the psychological determinants in the \textbf{human user}
    (see~\cite{epley_seeing_2007}),

@@

When taking into account these multiple factors, a logical consequence seems to
be that anthropomorphism is not either there or not there but that it arises in
different levels, may they be distinct, having multiple layers
\citep{persson_anthropomorphism_2000}. or a continuum being describes in low and
high anthropomorphism \citep{ruijten_introducing_2014}. For instance,

Moreover, we propose that anthropomorphism shows not only variances determined
by these four main factors but that -- the fourth factor implies this --
anthropomorphism is likely to evolve over time with continued interaction
between the human and the non-human agent.

To facilitate things, we introduce the notation \Ant[R,H,C,t] to bring these
multiple factors together. To take into account dynamics over time, we suggest
that anthropomorphism is a function of a vector of robot's traits $R =
\{r_0,...,r_n\}$, a vector human determinants $H = \{h_0,...,h_n\}$,
characteristics of the interaction context $C = \{c_0,...,c_n\}$, and the
duration / history of the interaction itself, $t$.

@@

% background shading
\path[fill=green!20] (0,0) rectangle (0.6,5.5);
\path[fill=blue!15] (0.6,0) rectangle (5.2,5.5);
\path[fill=orange!20] (5.2,0) rectangle (9.8,5.5);
\draw(0,5.5) node[anchor=south west, rotate=10] {\scriptsize \sc Initialization};
\draw(2.5,5.5) node[anchor=south west, rotate=10] {\scriptsize \sc Familiarization};
\draw(6.5,5.5) node[anchor=south west, rotate=10] {\scriptsize \sc Stabilization};
% horizontal axis
\draw[->] (0,0) -- (10,0) node[anchor=north] {$t$};
\draw(5,-0.1) node[anchor=north] {\scriptsize Duration of interaction};

@@

% vertical axis - end
\draw[->] (9.8,0) -- (9.8,2) node[anchor=east] {};
\draw (9.8, 0.8) node[anchor=west] {\SLA};

\draw[<-] (0.65,5) -- (0.8,5.2) node[anchor=east] {};
\draw (0.9,5.3) node[anchor=west] {\tiny \it novelty peak};

\draw[dashed] (0.6, 0) -- (0.6,5.1);
\draw (0.6,0) node[anchor=north] {$t_{max}$};

\draw[dotted] (0, 5) -- (6.2,5);
\draw[dotted] (0, 3) -- (6.2,3);
\draw[<->] (6.1,3) -- (6.1,5) node[sloped, above, midway] {$\Delta_{a}$};

\draw[<-] (1.7,4.4) -- (3.2,4.1) node[anchor=east] {};
\draw[<-] (2.85,2.6) -- (3.2,4.1) node[anchor=east] {};
\draw (3.2,4.1) node[anchor=west, align=center] {\tiny \it disruptive\\ \tiny
\it behaviors};
%%%%%
%% CURVES
%%%%
\begin{scope}[yscale=-1,shift={(-0.125,-0.4)}]

    \draw[ultra thick] svg[scale=1cm] "M 0.125,-2.5619582 c 0.0200733,-1.1573591
    0.33954391,-1.9982333 0.57940683,-2.0013597 0.13691605,0 0.25550329,0.052908
    0.37497897,0.1661788 0.119476,0.1130305 0.2396954,0.2786079
    0.37944,0.4790565 0.069872,0.099803 0.1021612,-0.2231749
    0.1792538,-0.2238964 0.099707,0 0.013454,0.4985362 0.3173383,0.9139831
    0.1911917,0.2612926 0.3444354,0.2578391 0.6711932,0.4894291
    0.2058265,0.1458798 0.1571532,0.9762306 0.2615031,0.9702183 0.097182,0
    0.083295,-0.3336208 0.4330446,-0.1199242 0.4306401,0.2631203
    0.216699,0.1355642 0.4296663,0.2647074 2.0858202,1.13836043
    4.34683,1.41300027 6.3741749,1.43753027";

    \draw[dashed] svg[scale=1cm] "M 0.125,-0.1952939 c 0.0553601,-3.1917862
    0.33954391,-4.3648976 0.57940683,-4.368024 0.13691605,0 0.25550329,0.052908
    0.37497897,0.1661788 m 2.6714393,2.7735737 C 4.7653984,-0.85050947
    6.5955093,0.35382293 10.125,0.4062498";

    \draw[dashed] svg[scale=1cm] "M 0.125,-3.6923825 C 0.1102888,-4.0297903
    0.46454391,-4.5601915 0.70440683,-4.5633179 M 3.7508251,-1.6235654 C
    4.9579368,-0.95548345 8.1358175,-0.82609971 10.125,-0.79459549";

@@

\caption{The dynamics of anthropomorphism. We distinguish three main phases:
    \emph{initialization}, \emph{familiarization} and \emph{stabilization},
    preceded by a \emph{pre-interaction} phase. The pre-interaction phase
    is characterized through an \emph{initial capital of anthropomorphism} (ICA, noted \ICA).
    Once the interaction starts, the level of anthropomorphism increases due to
    the \emph{novelty effect}, and then decreases to reach a \emph{stabilized
    level of anthropomorphism} (SLA, noted \SLA). \ICA and \SLA may vary
    depending on the user (and his/her expectations), the robot and the context
    of interaction.  During the interaction, unpredicted behaviors of the robot
    (\emph{disruptive behaviors}) may lead to local increases or decreases of
    the level of anthropomorphic effects.}

@@

First, the \emph{initialization} phase. During this short phase (from a couple
of seconds to a couple of hours), we observe an increased level of
anthropomorphism, from an \emph{initial capital of anthropomorphism} \ICA
(detailed in the next section) to a peak of anthropomorphic manifestations
\AntMax that corresponds to the maximum of the \emph{novelty effect}.
Section~\ref{sec:initialization} details this first phase.

The second phase, \emph{familiarization}, lasts longer (up to several days) and
models the process of the human getting acquainted to the robot: by observation
and interaction, the human builds a model of the robot's behavior that allows
him/her to predict the robot's actions. We observe a decrease of
anthropomorphic effects during this phase, that we explain by the acquired
ability to predict the behavior of the robot: the initial apparent behavioral
complexity vanishes, and the robot is considered more and more as a tool.
Section~\ref{sec:familiarization} discusses the second phase.

The last phase is the \emph{stabilization} phase. The level of anthropomorphic
effects tends to stabilize over a longer time, to reach a \emph{stabilized
level of anthropomorphism} \SLA. \SLA may be null (no anthropomorphic
effects observed anymore), but it may also remain at a higher level.  This
third phase, as well as the \emph{stabilized level of anthropomorphism}, are
discussed in section~\ref{sec:stabilization}.

\paragraph{Pre-interaction phase and Initial Capital of Anthropomorphism (ICA)\\}
\label{sec:ica}

The \emph{initial capital of anthropomorphism} \ICA describes the initial
potential for the robot to be anthropomorphized by the human user in a given
situation.  This potential depends on several factors, introduced in
section~\ref{sec:multi-factors}, see also Figure \ref{fig:factors}. It has been
shown, for instance, that some \textit{people} tend to anthropomorphize more
than others, that some \textit{situations} induce anthropomorphism more than
others, that \textit{children} tend to anthropomorphize more than adults, and
that some \textit{cultures} are notorious for their anthropomorphic religions
and worldviews~\citep{epley_when_2008}. Also the shape and design of the robot
play a role, as well as the context in which the interaction takes place. Our
model of anthropomorphism represents these factors as a compound metric \ICA. As
such, \ICA describes the level of anthropomorphic projections just prior the
first (real or imagined) interaction with a robot. In this stage of
pre-interaction, people form \textbf{initial expectations} toward the robot and
imagine how they will use it / interact with it.

Taking up on our suggestion to have four main factors that impact on
anthropomorphism, we propose that the initial capital of anthropomorphism can be
build \emph{a priori}.  As the \ICA marks the very moment before the first real
or imagined contact of the human with the robot, the \textit{interaction
history} between the two players is empty; consequently the time factor $t$ is
not taken into account when building the \ICA.  The specific \ICA value reflects
how far a robot is likely to be anthropomorphized, based on the following three
factors:

    \item ~\emph{Human factor} $S$: The \textbf{personality} and individual
        traits of the human user: Psychological characteristics / determinants
        that influence a person's tendency to anthropomorphize
        artifacts~\citep{epley_seeing_2007}. Other individual traits and
        demographic aspects are comprised (\eg age, gender, cultural background,
        professional background).

    \item ~\emph{Robot factor} $R$: The robot's \textbf{design} and how it
        appears to the human user. Characteristics of the robot's form,
        behavior, and interaction modalities (anthropomorphic
        design)~\citep{fong_survey_2003}.

    \item ~\emph{Context factor} $C$: The real or imagined
        \textbf{purpose} of the robot, including the situational context in
        which it is used, as well as the task context and role in which the
        robot is used / experienced (environmental
        context)~\citep{joosse_what_2013}.

For \textbf{personality}, we suggest to apply the psychological \textit{3-factor
theory} of anthropomorphism by \citet{epley_seeing_2007}. For instance, children
generally tend to anthropomorphize objects more than adults. Also, a person who
lacks social connection is said to be more likely to anthropomorphize. Both
aspects would increase \ICA. This means that, one and the same robot used by a
different user, can lead to a different \ICA.

For \textbf{design}, we understand that a robot that follows an anthropomorphic
design up to a certain acceptable level (\eg Nao) leads to a higher \ICA than a
rather functional robot (\eg Roomba). Also, a robot that is able to display
facial expressions would increase the \ICA. Quantifying the \emph{level} of
anthropomorphic design in a robot, as suggested for instance by
\citet{fong_survey_2003}, can be difficult.  Nevertheless, the important aspect
is how the robot appears to the user, \eg human-like or machine-like, acceptable
or not-acceptable.\footnote{It becomes clear that the \textit{robot factor} can
not be fully separated from the \textit{human factor}. The reason for this is
that the message that design tries to make, always lies in the eye of the
beholder. No matter how much effort a designer spends in making the ``message''
of the artifact's design clear, there is always some scope and creative freedom
in how the design is interpreted by the beholder.}

By taking the \textbf{purpose} of a robot into account, we suggest that the real
or imagined application context of the robot, or in which situation the
interaction occurs, impacts how far the robot will be attributed human-like
characteristics. We draw on findings such as presented in
\citet{joosse_what_2013}. The authors showed for instance that when the same
robot (Nao) is used in a different task context (cleaning task \emph{vs.} tour
guide), users ascribe different personalities to the robot. In general, a robot
which is imagined in a social, entertaining or playful context leads to a higher
\ICA than a robot which is used for a routine or focused task (security, rescue,
etc.). This was also found in an analysis of anthropomorphic language in online
forum discussions about the AIBO, Roomba, and iPad
\citep{fink_anthropomorphic_2012}. Regardless of the conversation topic, more
anthropomorphic language was found in the AIBO forum. However, this was not only
due to the zoomorphic (pet-like) shape of AIBO but also due to its  playful
purpose. The idea that the purpose of a robot are crucial factors, also receives
support from Goetz \& Kiesler's work that revealed that people prefer a serious
robot for serious tasks and a less serious robot for more playful
tasks~\citep{goetz_cooperation_2002, goetz_matching_2003}.  However, what about
artificial agents that seem to have no purpose? \cite{kaplan_free_2000}
discusses the role of uselessness in the design of robots. He argues that
artificial pets such as the AIBO, have no real purpose, in a sense that they do
not provide any kind of service, or, in other words \textit{``they are not doing
what you tell them to do''} and \textit{``they might refuse the order of its
owner''}. It is this very aspect that might increase people's tendency to
anthropomorphize the robot. According to Kaplan, it is in our daily use
language, that we tend to attribute intentions to devices that are not doing
their job well. For instance, we do not develop any kind of relationship with
our computer, washing machine or TV set when they work properly. The author
argues that it is only when they start dis-functioning, when they show that they
can act differently that what we ordered, that we are ready to give them some
kind of ``intentionality''.

Also, we suggest that the environmental context in which people experience and
interact with the robot impacts the \ICA. For instance, several friends
interacting simultaneously with the robot might lead to increased \ICA for each
of the users, due to increased human-human social interactions (the robot might
be perceived to be part of the social interaction, and in turn attributed
human-like qualities~\citep{baxter2013do}).

\subsubsection{Initialization Phase\\}
\label{sec:initialization}

%\paragraph{The Novelty Effect\\}
%\label{sec:noveltyeffect}

We assume that the short phase of initialization is strongly affected by the
so-called novelty effect.  Novelty effects have been described with interactive
technologies in general \citep{rogers_diffusion_1995} and also in several
long-term HRI studies in workplaces
\citep{huttenrauch_fetch-and-carry_2003,mutlu_robots_2008}, in schools and
public places
\citep{gockley_designing_2005,kanda_interactive_2004,kanda_communication_2005},
or elder care centers \citep{sabelli_conversational_2011}. Further, also with
domestic robots it has been found that people's interest, engagement, and
fascination with the robot decreased over time
\citep{fink_living_2013,fernaeus_how_2010,sung_robots_2009}.

To illustrate how the novelty of a robot can evoke anthropomorphic effects, we
present an example from our long-term ethnographic study with Roomba vacuum
cleaning robots in households \citep{fink_living_2013}. The observed scenario
describes how an elderly woman (E., 71 years, living on her own with a little
dog in a small row house) was setting up and using the robot for the first time
during about 10 minutes.

At several occasions, while watching how the robot moves around and cleans, E.
makes compliments to the robot: \emph{``Nice!''} or \emph{``Good job!''} or
\emph{``Bravo!''}. Once, when the Roomba turns right in front of some crumbs on
the kitchen floor and moves away without cleaning them up, E. says:
\emph{``There you forgot something, what about over there? Look, there you still
have something to clean up!''}, and she points at where she wants the Roomba to
go. Further, she uses deictic gestures and tells the robot: \emph{``No, go this
way! Come over here!''}. The woman keeps observing the robot and wonders about
its intentions, asking it directly: \emph{``Why do you leave the kitchen? There
is still something there on the fringe!''}. E. turns to the experimenter, and
tells her: \emph{``She [the robot] loves the hall!''}.

Besides this behavior of generally anthropomorphizing the robot, we observe the
following concrete instances as anthropomorphic effects in this scenario:
talking directly to the robot, or asking it questions (though it doesn't
recognize speech); giving it nick-names or ascribing a gender, praising the
robot, using pointing gestures toward the robot, or trying to direct it around
with commands, and attributing a personal preference / intention to the robot
(\emph{``she loves the hall!''}).

During the following visits at the woman's place, E. did not talk to the robot
anymore, and generally behaved in a less anthropomorphizing way. The observed
effects were likely due to the novelty of the robot, and the fact that E. was
not familiar with the robot and its behavior.  As
\cite{persson_anthropomorphism_2000} names it: \textit{``Anthropomorphism is a
way of making sense of complex behavior in the world around us.''} Thus, as soon
as this complexity is reduced, also anthropomorphism might be reduced.

\subsubsection{Familiarization Phase\\}
\label{sec:familiarization}

In this context, \emph{familiarization} means \emph{to get acquainted with}.
It is not to be confused with the \emph{familiarity thesis} that relates to the
projection of already known cognitive models onto the robot. We discuss this other
aspect in section~\ref{sec:cognition-neuroscience}.

The \emph{familiarization phase} starts at the peak of the \emph{novelty
effect}: anthropomorphic effects are at their maximum, the user thinks (s)he is
potentially facing an agent aiming at ``human-level intelligence'' (to take
McCarthy's words). This is a transient state that quickly vanishes, and the
projected anthropomorphism then starts to decrease while the human observes and
recognizes that the behavior of the robot is generally \emph{predictable} and
possibly \emph{deceptive} (for instance, it talks but does not understand when
we talk; it has eyes but does not recognize everyday objects we show; \etc).

\paragraph{Disruptive behaviors\\}

By \emph{disruptive behaviors}, we mean any behavior exhibited by the robot that
appears unexpected to the user: for instance, a robot may usually follow always
the same route to go from one place in a house to the other, but suddenly it
might change the route. The actual underlying reason may span from a bug to the
detection of a new obstacle, but as long as this reason is not immediately
intelligible to the user, the behavior counts as \emph{disruptive}. (We will
come back to the role of disruptive behaviors later, in Section
\ref{sec:disruptive}.)

As illustrated in Figure~\ref{fig:dynamics}, our model represents
\emph{disruptive behaviors} as either local increases or local decreases of
anthropomorphic effects in the familiarization phase (and at a lesser extend, in
the stabilization phase): because such behaviors are unexpected, a human
observer may interpret them either as the result of a richer deliberative
process, which in turn leads to the supposition of complex cognitive skills, or
conversely as a failure, reminding the human that the robot is ``just a
machine''. Section~\ref{sec:disruptive} discusses more in depth the impact of
the user's perception of disruptive behaviors, and their cognitive correlates.

To illustrate how a disruptive behavior of a robot can evoke anthropomorphic
effects, we present again an example from our long-term ethnographic study with
Roomba vacuum cleaning robots in households \citep{fink_living_2013}. The
observed scenario describes how a family (young parents, 26 and 28 years old,
with 3 little children, 5 and 4 years, and 6 months old, living in a house with
garden on the countryside) were using the robot during the first week.

The father (M.) always lets Roomba clean up the ground floor after the family
finishes dinner and they all go upstairs to put the children to bed. When the
robot would have finished its cleaning cycle, the parents would later just find
it docked at its charging station. It used to be like this every evening.
However, one day when the father comes downstairs, Roomba is not at its station.
He says: \emph{``Hey, Roomba, where are you? Why did you not want to got to
sleep last night?''}. After some searching he found it stuck under the sofa and
wondered why this had happened.

What is the observed anthropomorphic effect in this scenario? Usually, the
father did not talk to the robot or praise it. However, now that the robot did
something unexpected (it was not at its usual place) he started wondering about
its ``intention'' and directly asked the robot for its reasons to not
\textit{``go to sleep''}.

\subsubsection{Stabilization Phase\\}
\label{sec:stabilization}

The last phase is the \emph{stabilization} phase. The level of anthropomorphic
effects tends to stabilize over a longer time, to reach a \emph{stabilized
level of anthropomorphism}, noted \SLA.

If, after the familiarization phase, no anthropomorphic effects are observed
anymore, $\sla = 0$. This can be interpreted as the user interacting with the
robot in a routine way, without projecting anymore human-like traits on the
robot.

\paragraph{Stabilized Level of Anthropomorphism\\}

The \emph{Stabilized Level of Anthropomorphism} describes the long-term lasting
level of anthropomorphism.  We proposed that \ICA is built on three factors:
user's \emph{personality}, robot's \emph{design} and interaction \emph{purpose}
(or \emph{interaction context}). The user's personality and the context of use
do also influence \SLA.  In particular, it appears that the user's level of
acquaintance with technologies plays an important role in long-term tendency to
anthropomorphize~\citep{fink_living_2013} (people more familiar with technology
understand, and hence predict, better the behavior of the robot, which in turn
leads them more frequently to ultimately consider the robot as a simple tool).

>>>>

\usepackage{graphicx}
\graphicspath{{figs/}}
\usepackage{tikz}
\usetikzlibrary{shapes}
\usetikzlibrary{svg.path}
%
\usepackage{amsmath} 
\usepackage{mathptmx}      % use Times fonts if available on your TeX system
%
\usepackage{color, soul}
\usepackage{url}
\usepackage{multirow}
\usepackage{array}
\usepackage{fixltx2e}
\usepackage{textcomp}
\usepackage{booktabs}

@@

@@

The remainder of this article is organized as follows.  According to the three
perspectives that we take, we first (in Section \ref{sec:anthropomorphism}) aim
to establish a common ground in the understanding of anthropomorphism. We review
literature from several relevant domains and integrate the different
explanations into an understanding of anthropomorphism as a multi-factor
phenomenon.  Based on this, Section~\ref{sec:our-ideas} proposes a new
conceptual model of the \textit{Dynamics of Anthropomorphism}. This model
describes the variances and dynamics of people's tendency to anthropomorphize an
artificial agent, such as a robot.  Then Section
\ref{sec:cognition-neuroscience} presents a cognitive and neuroscience
perspective on anthropomorphism to consolidate our considerations and the
proposed model.  Section \ref{sec:measuring} looks at how these models can be
operationalized, and in particular, what are the tools at hand to measure
anthropomorphism. Finally, Section \ref{sec:conclusion} concludes this article
and offers starting points for further research on anthropomorphism and the
effect of the human-like design approach in robotics.

@@

\section{The Dimensions of Anthropomorphism}
\label{sec:anthropomorphism}

Literature on anthropomorphism is quite diverse. There exist different and
partly conflicting definitions of the term anthropomorphism, and it still
appears controversial that humans tend to \textit{``see human
in the non- human''} \citep{epley_seeing_2007}.

As anthropomorphism arises in an interaction between a human and a non-human
agent~\citep{persson_anthropomorphism_2000}, one can also take two main
\textit{perspectives} when trying to explain why humans anthropomorphize~\citep{lee_human_2005}: an artifact-centered perspective and a human-centered
perspective (the grey triangles in Figure \ref{fig:anthropofig}).

In the following we will only briefly present the artifact-centered perspective
to explain anthropomorphism. A review of anthropomorphic design of robots can
also be found in~\cite{fink_anthropomorphism_2012}. Then, we provide a more
detailed review and synthesis of literature that applies a human-centered
perspective to explain anthropomorphism. We aim here at providing an exhaustive
picture of this phenomenon, by summarizing theories from developmental
psychology, cognitive science, neuroscience, and social psychology.

@@

\begin{figure}
    \centering
    \includegraphics[width=0.75\columnwidth]{anthropo.pdf}
    \caption{Anthropomorphism is a psychological phenomenon which arises in a
        social context, during an interaction between a human agent and a
        non-human agent. One can take two main perspectives when trying to
        explain anthropomorphism: a human-centered and an artifact-centered
        perspective (grey triangles). Several external factors influence the
        anthropomorphism phenomenon. Person- and artifact-related characteristics,
        but also related to the context of use as well as the actual time spent
        interacting with the robot (interaction history).}

@@

@@

\item ~the psychological determinants in the \textbf{human
    user}~\citep{epley_seeing_2007},

@@

As such, we use in the remaining of the article the notation \Ant[R,H,C,t] to
represent anthropomorphism as a function of the robot's traits $R$, the human
determinants $H$, the characteristics of the interaction context $C$ and the
history of the interaction itself, $t$. This last factor, time, is the focus of
the next section.

@@

% background shading
\path[fill=green!20] (0,0) rectangle (1.1,5.5);
\path[fill=blue!15] (1.1,0) rectangle (5.2,5.5);
\path[fill=orange!20] (5.2,0) rectangle (9.8,5.5);
\draw(0.5,5.5) node[anchor=south west, rotate=10] {\scriptsize \sc Initialization};
\draw(3,5.5) node[anchor=south west, rotate=10] {\scriptsize \sc Familiarization};
\draw(6.5,5.5) node[anchor=south west, rotate=10] {\scriptsize \sc Stabilization};
% horizontal axis
\draw[->] (0,0) -- (10,0) node[anchor=north] {$t$};
\draw(5,-0.1) node[anchor=north] {\scriptsize Interaction history};

@@

% vertical axis - end
\draw[->] (9.8,0) -- (9.8,4) node[anchor=east] {};
\draw (9.8, 0.8) node[anchor=west] {\SLA};

\draw[<-] (1.15,5.1) -- (1.30,5.3) node[anchor=east] {};
\draw (1.35,5.4) node[anchor=west] {\tiny \it novelty peak};

\draw[dashed] (1.1, 0) -- (1.1,5.1);
\draw (1.1,0) node[anchor=north] {$t_{max}$};

%\draw[dotted] (0, 5) -- (6.2,5);
%\draw[dotted] (0, 3) -- (6.2,3);
%\draw[<->] (6.1,3) -- (6.1,5) node[sloped, above, midway] {$\Delta_{a}$};

\draw[<-] (2.2,4.4) -- (3.7,4.1) node[anchor=east] {};
\draw[<-] (3.3,2.6) -- (3.7,4.1) node[anchor=east] {};
\draw (3.8,4.1) node[anchor=west, align=center] {\tiny \it disruptive\\ \tiny
\it behaviors};
%%%%%
%% CURVES
%%%%
\begin{scope}[yscale=-1,shift={(-0.125,-0.46)}]
    \draw[ultra thick, green!70!red!50, dashed] svg[scale=1cm] "M 0.114,-1.1 C 0.733,-2.79 1.88,-2.23
    2.51,-1.76 3.34,-1.14 3.94,0.02 10.1,0";

    \draw[ultra thick, blue!70!red!50, dashed] svg[scale=1cm] "M 0.125,-3.21 C 0.549,-4.91 2.05,-4.42
    2.59,-3.83 3.57,-2.74 5.93,-1.51 10.1,-1.53";

    \draw[ultra thick, red!70, dashed] svg[scale=1cm] "M 0.12500001,-2.7222489 C 0.70533638,-5.3401235 1.9431386,-3.7307246 2.608741,-3.3299187 c 1.3619528,0.8201274 3.3166657,1.2251393 7.515191,1.2014756";

    \draw[ultra thick] svg[scale=1cm] "M
    0.125,-1.622832 c 0.19887378,-2.35086 0.74751962,-2.9373641 1.0765913,-2.9404859 0.3290717,-0.00312 0.5370891,0.383599 0.754419,0.6452353 0.069872,0.099803 0.1021612,-0.2231749 0.1792538,-0.2238964 0.099707,0 0.013454,0.4985362 0.3173383,0.9139831 0.1911917,0.2612926 0.3444354,0.2578391 0.6711932,0.4894291 0.2058265,0.1458798 0.1571532,0.9762306 0.2615031,0.9702183 0.097182,0 0.083295,-0.3336208 0.4330446,-0.1199242 0.4306401,0.2631203 0.216699,0.1355642 0.4296663,0.2647074 2.0858202,1.13836043 3.8496455,1.41300027 5.8769904,1.43753027";

@@

\caption{\textbf{Conceptual model of the dynamics of anthropomorphism}. We
    distinguish three main phases: \emph{initialization}, \emph{familiarization}
    and \emph{stabilization}, preceded by a \emph{pre-interaction} phase. The
    pre-interaction phase is characterized through an \emph{initial capital of
    anthropomorphism} (ICA, noted \ICA).  Once the interaction starts, the level
    of anthropomorphism increases due to the \emph{novelty effect}, and then
    decreases to reach a \emph{stabilized level of anthropomorphism} (SLA, noted
    \SLA). \ICA and \SLA may vary depending on the user (and his/her
    expectations), the robot and the context of interaction.  During the
    interaction, unpredicted behaviors of the robot (\emph{disruptive
    behaviors}) may lead to local increases or decreases of the level of
    anthropomorphic effects. The duration of these phases may vary and the
    overall shape of the curve is unique to each human interacting with a
    specific robot in a given context, as depicted by the alternative curves.}

@@

\subsubsection{Initialization Phase\\}

The \textit{Initialization Phase} starts with the so-called \textit{Initial
Capital of Anthropomorphism} ($ICA$), which corresponds to \ICA (value of \Ant
at time point $0$). This value considers the multiple factors presented in
Section~\ref{sec:multiple-factors}, thus it is composed by factors of the human
$H$, the robot $R$, and the context of use $C$. As such, \ICA describes the
value of anthropomorphic effects that can be expected to occur during the first
contact between the human and the robot in a given situation. During this first
phase, pre-conceptions and initial expectations as well as previous experience
of the user with other similar types of robots and similar contexts can have a
strong impact and influence the general tendency to anthropomorphize.

During the \textit{Initialization Phase} we expect an increasing value of $A$
that may be partly due to the so-called \textit{effects of novelty} of the
robot. Such effects have been described with interactive technologies in general
\citep{rogers_diffusion_1995}, and also in several long-term HRI studies in
workplaces \citep{huttenrauch_fetch-and-carry_2003,mutlu_robots_2008}, in
schools and public places
\citep{gockley_designing_2005,kanda_communication_2005,kanda_interactive_2004},
and elder care centers \citep{sabelli_conversational_2011}. Also long-term
studies of domestic robots have shown similar effects and found that after some
time people's interest, engagement, and fascination of the robot decreased
\citep{sung_robots_2009,sung_domestic_2010,fernaeus_how_2010,fink_living_2013}.

What type of anthropomorphic effects are typical during the
\textit{Initialization Phase}? In our own work, we observed that users
spontaneously tend to talk directly to the robot (though the user knows that the
system is not be able to recognize speech), for instance, by greeting the robot,
asking questions to it, giving it nick-names or ascribing a gender, praising the
robot, commenting on its performance in general, and giving commands. Also, the
use of point gestures and waving at the robot were fairly common among
participants of our studies, as well as ascribing gender and some personal
preferences or intention to the robot. (Note that we did not even use an
anthropomorphic robot but a Roomba vacuum cleaning robot.)

In general, it appears that anthropomorphism during the \textit{Initialization
Phase} is a commonly observed phenomenon. But we would like to stress again that
most of these effects are likely to fade out over time.

\subsubsection{Familiarization Phase\\}

In the \textit{Familiarization Phase} the user gets acquainted and familiar with
the robot and with using the robot, and hence the tendency to anthropomorpize
decreases. This phase typically starts as the local maximum value of
anthropomorphic effects that are enhanced by the novelty of the robot start to
wear off. The user may think that he/she faces an agent that aims at
``human-level intelligence'' (to take McCarthy's words). This is a transient
state that quickly vanishes as the user recognized that the behavior of the
robot is generally \textit{predictable} and possibly \textit{deceptive}. For
instance, the robot produces verbal utterances (it ``talks'') but it does not
understand when we talk; or the robot is endowed with eyes but does not
recognize the objects we show it.

\paragraph{Disruptive behaviors.}

We call \emph{disruptive behaviors} any behavior exhibited by the robot hat
appears unexpected to the user: for instance, a robot may usually follow always
the same route to go from one place in a house to the other, but suddenly it
might change the route. The actual underlying reason may span from a bug to the
detection of a new obstacle, but as long as this reason is not immediately
intelligible to the user, the behavior counts as \emph{disruptive}. We further
discuss the role of disruptive behaviors in cognitive terms in Section
\ref{sec:disruptive}.

\subsubsection{Stabilization Phase\\}

During the \textit{Stabilization Phase} the value of $A$ tends to stabilize over
a longer time, to reach a value that we call the \textit{Stabilized Level of
Anthropomorphism} ($SLA$), which corresponds to \SLA. The value that the
$SLA$ approaches is unique for each user interacting with a specific robot in a
given situation. \SLA may approach $0$, when no anthropomorphic effects
are observed anymore, for instance, when the user is interacting with the robot
in a routine way and he/she does not project any human-like traits unto the
robot\footnote{The robot may have become \textit{``invisible in use''}
\citep{takayama_toward_2011}.} As such, the $SLA$ stands for the long-term
lasting value of anthropomorphism.  We proposed that \SLA (the value
corresponding to the $ICA$) is built on multiple factors (see
Section~\ref{sec:multiple-factors}) and similar \SLA is based on these
factors, however taking into account the \textit{interaction history} $t$ that
the user has with the robot.  Note that the $ICA$ and the $SLA$ are generally
not correlated. A person may initially have a high potential of
anthropomorphizing (high \ICA), for instance a highly anthropomorphic robot,
but subsequently get disappointed by the actual capabilities of it and stop
explaining it in terms of human-like characteristics (low \SLA). Another
user may however create lasting affective bonds with the same robot and may
continue to anthropomorphize the robot (high \SLA).
